---
title: "MGMT gene methylation: TCGA data analysis"
author: "O. Celiku"
date: '`r format(Sys.time(), "%b %d %X %Y")`'
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
## for high resolution set dpi to 300
knitr::opts_chunk$set(echo = FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=100)
data_dir <- file.path("data")
```

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(kableExtra)
library(ggpubr)
library(viridis)
# color palette
palette <- viridis::turbo(2, begin =0.15, end = 0.9, direction = 1)
```

The input files were generated based ond downloaded files from Xena browser and filtering for samples that had relevant information:

* mRNA expression, filtered to MGMT only: tcga_RSEM_Hugo_norm_count.gz": mRNA expression
* Methylation, filtered to probes associated with MGMT (based on Illumina Methylation 450k array probe annotations obtained using methylationARrayAnalysis library): jhu-usc.edu_PANCAN_HumanMethylation450.betaValue_whitelisted.tsv.xena
* Phenotype: TCGA_phenotype_denseDataOnlyDownload.tsv.gz
* Mutation: mc3.v0.2.8.PUBLIC.nonsilentGene.xena.gz
 
```{r}
meth <- read_csv(file.path(data_dir, "TCGA_MGMT_meth.csv"))
pheno <- read_csv(file.path(data_dir, "TCGA_pheno_MGMT_mRNA.csv"))
```

### Descriptive landscape of pan-cancer, including glioma

First some overview of the landscape using most varied samples for each cancer type; showing here for top 50 most varied samples (excluding those that fall outside 95% quantile of sd), roughly grouping the probes by region with any probes annotated as promoter considered as promoter (despite CpG island status) and the rest as body. 

```{r, fig.width=8, fig.height=1.8}
topn <- 50
library(circlize)
col_fun = colorRamp2(c(0, 0.5, 1), c(palette[1], "white", palette[2]))

m <- as.matrix(meth[,pheno$sample])
rownames(m) <-meth$probe

m.ann <- meth[,1:5] %>% mutate(region=ifelse(promoter==TRUE, "promoter","body"))
m.sd <- apply(m,2,function(x){sd(x,na.rm=TRUE)})
m.sd <- bind_cols(sample=colnames(m),sd=m.sd, type=pheno$study)
m.sd <- m.sd %>% group_by(type) %>% 
  dplyr::summarize(q95 = quantile(sd, 0.95), sd=sd, sample=sample) %>%
  filter(sd <= q95) %>% top_n(topn, sd) 

m.values <- m[,m.sd$sample]
m.body <- m.values[m.ann%>% filter(region=="body")%>%dplyr::pull(probe),]
m.promoter <- m.values[m.ann%>% filter(region=="promoter")%>%dplyr::pull(probe),]
p.body <- Heatmap(m.body, column_split = m.sd$type, 
        show_row_names = FALSE, column_title_rot = 90,
        show_column_names = FALSE, column_title_gp = gpar(fontsize = 8), heatmap_legend_param = 
          list(title="body\nBeta values"), col=col_fun)
p.promoter <- Heatmap(m.promoter, column_split = m.sd$type, 
        show_row_names = FALSE, 
        show_column_names = FALSE, column_title_gp = gpar(fontsize = 8), column_title_rot = 90, heatmap_legend_param = list(title="promoter\nBeta values"), col=col_fun)
```

```{r, r, fig.width=8, fig.height=1.3}
p.promoter
```

```{r, fig.width=8, fig.height=1.5}
p.body
```

### Cohort specific associations of region methylation with mRNA

Here we compute the aggregated methylation levels per region per sample first aggregating probes by smaller regulatory regions and then promoter versus body: 

* For promoter values we aggregate the values that are promoter associated per annotations, as well as fall on CpG islands.
* For body we use all values that are not promoter associated.

```{r, fig.width=10, fig.height=3, eval=TRUE}
long <- meth %>% filter((promoter==FALSE) | (island=="Island" & promoter==TRUE)) %>% pivot_longer(!c(probe, island, group, enhancer, promoter), names_to = "sample", values_to = "value") %>% group_by(sample, island, group, enhancer, promoter) %>% dplyr::summarize(methylation=mean(value, na.rm = TRUE)) %>% ungroup()

long <- long %>% left_join(pheno) %>% transmute(sample, type=study, cohort, region=ifelse(promoter==TRUE, "promoter","body"), methylation, mRNA=MGMT_mRNA) %>% group_by(sample, type, cohort, region,  mRNA) %>% dplyr::summarize(methylation = mean(methylation, na.rm=TRUE)) %>%
  filter(!is.na(type))

wide <- long %>% pivot_wider(names_from=region, values_from=methylation)
```

We consider pan-cancer excluding glioma cohort separately; as well as the glioma cohorts split by IDH mutation status.

First correlation analyses, as well as linear regression of association of mRNA expression with promoter and body methylation, considering models with and without interaction. 

#### Non-glioma

```{r, fig.width=3, fig.height=2.7}
p <- ggplot(wide %>% filter(cohort=="non-glioma"),aes(body, promoter)) +
  geom_point() +
  ggtitle("Non-glioma\npromoter body correlation") +
  stat_smooth(method = "lm", level=0.95) + theme_minimal() 
p + stat_cor(label.x = 0.2, label.y = 0.75, p.accuracy = 0.001, r.digits=2)
```
```{r,fig.width=5, fig.height=2.7}
p <- ggplot(long%>% filter(cohort=="non-glioma") %>%
         mutate(region=factor(region, levels=c("promoter", "body"))),aes(mRNA, methylation)) +
  geom_point(aes(group=region, color=region)) +
  stat_smooth(aes(group=region), method = "lm", level=0.95) + theme_minimal()  +
  facet_wrap(~region) + 
  ggtitle("Non-glioma\nmRNA methylation correlation") +
  scale_color_manual(values=palette) +
  theme(strip.text.x = element_blank())
p + stat_cor(aes(group=region), label.x = 0.3, label.y = 0.99, r.digits=2, p.accuracy = 0.001)
```

Regression model without interaction; non-glioma:

```{r}
model <- lm(mRNA ~ promoter+body, data = wide %>% filter(cohort=="non-glioma"))
summary(model)
```

Regression model with interaction; non-glioma:

```{r}
model <- lm(mRNA ~ promoter*body, data = wide %>% filter(cohort=="non-glioma"))
summary(model)
```

#### Glioma IDH-wild-type

```{r, fig.width=3, fig.height=2.7}
p <- ggplot(wide %>% filter(cohort=="glioma IDHWT"),aes(body, promoter)) +
  geom_point() +
  ggtitle("IDH-wt glioma\npromoter body correlation") +
  stat_smooth(method = "lm", level=0.95) + theme_minimal() 
p + stat_cor(label.x = 0.7, label.y = 0.75, p.accuracy = 0.001, r.digits=2)
```
```{r,fig.width=5, fig.height=2.7}
p <- ggplot(long%>% filter(cohort=="glioma IDHWT") %>%
         mutate(region=factor(region, levels=c("promoter", "body"))),aes(mRNA, methylation)) +
  geom_point(aes(group=region, color=region)) +
  stat_smooth(aes(group=region), method = "lm", level=0.95) + theme_minimal()  +
  facet_wrap(~region) + 
  ggtitle("IDH-wt glioma\nmRNA methylation correlation") +
  scale_color_manual(values=palette) +
  theme(strip.text.x = element_blank())
p + stat_cor(aes(group=region), label.x = 6, label.y = 0.99, r.digits=2, p.accuracy = 0.001)
```

Regression model without interaction; IDH-wt glioma:

```{r}
model <- lm(mRNA ~ promoter+body, data = wide %>% filter(cohort=="glioma IDHWT"))
summary(model)
```

Regression model with interaction; IDH-wt glioma:

```{r}
model <- lm(mRNA ~ promoter*body, data = wide %>% filter(cohort=="glioma IDHWT"))
summary(model)
```

#### Glioma IDH-mutant

```{r, fig.width=3, fig.height=2.7}
p <- ggplot(wide %>% filter(cohort=="glioma IDHMut"),aes(body, promoter)) +
  geom_point() +
  ggtitle("IDH-mutant glioma\npromoter body correlation") +
  stat_smooth(method = "lm", level=0.95) + theme_minimal() 
p + stat_cor(label.x = 0.5, label.y = 0.75, p.accuracy = 0.001, r.digits=1)
```
```{r,fig.width=5, fig.height=2.7}
p <- ggplot(long%>% filter(cohort=="glioma IDHMut") %>%
         mutate(region=factor(region, levels=c("promoter", "body"))),aes(mRNA, methylation)) +
  geom_point(aes(group=region, color=region)) +
  stat_smooth(aes(group=region), method = "lm", level=0.95) + theme_minimal()  +
  facet_wrap(~region) + 
  ggtitle("IDH-mutant glioma\nmRNA methylation correlation") +
  scale_color_manual(values=palette) +
  theme(strip.text.x = element_blank())
p + stat_cor(aes(group=region), label.x = 5, label.y = 0.99, r.digits=2, p.accuracy = 0.001)
```

Regression model without interaction; IDH-mutant glioma:

```{r}
model <- lm(mRNA ~ promoter+body, data = wide %>% filter(cohort=="glioma IDHMut"))
summary(model)
```

Regression model with interaction; IDH-mutant glioma:

```{r}
model <- lm(mRNA ~ promoter*body, data = wide %>% filter(cohort=="glioma IDHMut"))
summary(model)
```

```{r}
sessionInfo()
```